---
title: "Data Imputation and Cleaning of Property Values"
output: html_notebook
---


```{r,echo=FALSE}
#0. read in data and library -----------------------------------------------------------
path <- "~/PetGit/ImputationClusterAnalysis/"
setwd(path)
library(data.table)
library(reshape2)
library(stringr)
library(Amelia)
library(Rcpp) 
library(DMwR2) # pkg impute missing values from median/mode of k nearest neighbors
library(ggplot2)
library(Rcpp)
library(plyr)
```



```{r}
# summary function for easy display of data
#>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>

line<-function(x){paste(replicate(x, "="), collapse = "")}

sum.funct<-function(my.df,i,j,n){
  print(summary(data.frame(my.df[,i:j])))
  print(line(n))}
```

```{r echo=FALSE, message=FALSE,echo=FALSE, warning=FALSE, message=FALSE}
# Import data
#>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
data.url<-"https://raw.githubusercontent.com/EGjohnson/ImputationClusterAnalysis/7959de2efcf1e264daf8dad2b604e6718d411409/Sales_new.txt"
#prop.sale <- fread(data.url,na.strings = c(""," ",NA,"NA"))
prop.sale<-fread("property_data/Sales_new.txt",na.strings = c(""," ",NA,"NA"))
prop.recs<-fread("property_data/Parcels_new.txt")
```

### 1. About The Two Data Sets Used in the Analysis
All records were imported from
http://hendersoncountync.org/gis/download_realproperty.html

#### Sales_new.txt 
This database file contains property sale records over the last decade. These records including information such as building size, year of the build, what the property was sold for and when. It contains `r dim(prop.sale)[1]` property sale records.
```{r, echo=FALSE}
head(prop.sale)
```


#### Parcels_new.txt 
This database file contains parcel records over the last decade. These records including information such as zoning, tax assessed value, neighborhood and acreage. It contains `r dim(prop.recs)[1]` parcel records.
```{r, echo=FALSE}
head(prop.recs)
```

Keep REID, FIRE_DISTRICT, LAND_CLASS,CALCULATED_ACRES,ZONING,ETJ,PHYADDR_CITY,PHYADDR_ZIP,TOTAL_LAND_VALUE_ASSESSED,TOTAL_BLDG_VALUE_ASSESSED,LAND_USE_VALUE,TOTAL_PROP_VALUE,COST_TOTAL_VALUE,NEIGHBORHOOD,LOCATION_ADDR



### Summary of Merged Data

Need to figure out what to do with data that says "other" in factors as well as outliers in numerical data.
```{r,echo=FALSE}
# Drop unwanted data
#>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
keep.recs.vec<-c("REID", "FIRE_DISTRICT", "LAND_CLASS","CALCULATED_ACRES","ZONING","ETJ","PHYADDR_CITY","PHYADDR_ZIP","TOTAL_LAND_VALUE_ASSESSED","TOTAL_BLDG_VALUE_ASSESSED","LAND_USE_VALUE","TOTAL_PROP_VALUE","COST_TOTAL_VALUE","NEIGHBORHOOD","LOCATION_ADDR")
prop.sale <- data.frame(subset(prop.sale, select = -c(SALE_PK,AUT_SNAPSHOT_DATE,VCS,DIST,SALE_PRICE_SOURCE,SEQ) ))
prop.recs<-data.frame(subset(prop.recs,select=keep.recs.vec))
```

```{r,echo=FALSE}
# Convert datatypes
#>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
factor.this<-c("SALE_TYPE","STORY_HEIGHT")
prop.sale[,c(3,6)]<-lapply(prop.sale[,c(3,6)], as.factor) #categorical variables to factor
prop.sale[,c(4,5,7)]<-lapply(prop.sale[,c(4,5,7)], as.numeric) #categorical variables to factor
prop.sale[,8]<-as.Date(prop.sale$SALE_DATE)
prop.sale$sale.years<-as.numeric((prop.sale$SALE_DATE-min(prop.sale$SALE_DATE)))/365

factor.this<-c(2,3,6,7,8,14)
numeric.this<-c(1,4,9,10,11,12,13)
character.this<-c(5,15)
prop.recs[,factor.this]<-lapply(prop.recs[,factor.this],as.factor)
prop.recs[,numeric.this]<-lapply(prop.recs[,numeric.this],as.numeric)
prop.recs[,character.this]<-lapply(prop.recs[,character.this],as.character)
```

```{r}
sum.funct(prop.sale,1,4,150)
sum.funct(prop.sale,5,9,150)
sum.funct(prop.recs,1,4,150)
sum.funct(prop.recs,5,dim(prop.recs)[2]/2,150)
sum.funct(prop.recs,dim(prop.recs)[2]/2,dim(prop.recs)[2],150)
```
```{r}
my.bins<-60
min.val<-100000
max.val<-300000
my.col<-c("PRICE","darkgreen")

prop.sale.sub<-subset(prop.sale,PRICE>min.val & PRICE<max.val)
ggplot(data=prop.sale.sub,aes_string(my.col[1])) + 
geom_histogram(bins=my.bins,fill=my.col[2])+xlim(c(min.val,max.val))
```

```{r}
saveRDS(prop.sale,"propsale.rds")
```

We can see in the below tabplot that (as is to be expected) we have missing values in YEAR_BUILT, STORY_HEIGHT AND BLDGE_SIZE for the LAND ONLY category in SALE_TYPE.
```{r Fig2, echo=TRUE, fig.height=5, fig.width=10}
#readRDS("propsale.rds)
library(tabplot)
tableplot(prop.sale,select=c(PRICE,SALE_TYPE,YEAR_BUILT,STORY_HEIGHT,BLDG_SIZE,SALE_DATE),sortCol=SALE_TYPE)
```

```{r Fig1, echo=TRUE, fig.height=8, fig.width=10}
tab.prop.dat<-subset(prop.sale,SALE_TYPE!="LAND ONLY")
tableplot(tab.prop.dat,select=c(PRICE,SALE_TYPE,YEAR_BUILT,STORY_HEIGHT,BLDG_SIZE,SALE_DATE),sortCol=BLDG_SIZE,from=0,to=100)
```

